---
title: "Isotope_Model"
author: "Amy Hessl"
date: "4/17/2020"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Learn exactly what ECHAM is
Read proxy-system model paper (Evans, Kevin, etc)
Schmidt
PRYSM paper
Steiger
Report back on that 

netcdf tutorials from Oregon (some functions are named different)
metadata is part of the data

Steps:
Go through
Do percentiles/quantiles
Then look at seasonal. Calc Oct-Feb on monthly marshall sam data. Create a factor - then use apply
After that, maybe adjust seasonal part to match what I did with SAM data





```{r}
library(ncdf4)
library(raster)
library(dplyr)
library(lubridate)
library(ggplot2)
library(lattice)
library(rnaturalearth)
library(rnaturalearthdata)
library(rasterVis)
library(viridis)
library(seas)
library(rgeos)
```

### Data Download from ECHAM5‐wiso (monthly, 1 degree grid)
https://zenodo.org/record/1249604#.XpnVbFNKhTZ

Add this to the data folder but DO NOT COMMIT BEFORE REMOVING THIS FILE!!!
It is too large to push to github, and it will be impossible to make any more pushes afterwards

Data sources, years, extent (could be variables in function call someday)
```{r}
netcdf.file <- "data/d18O_precip_mon_1871_2011.nc"

F_yr <- 1958
L_yr <- 2011
years <- seq(F_yr, L_yr)
ext <- extent(143.5, 148.6, -43.8, -39.5) #tas extent (lon_min, lon_max, lat_min, lat_max)
#ext <- extent(0, 180, -90, 0) #SH extent
```

Get a list of the variables to choose from and confirm that the time 
origin and units are appropriate......
```{r}
nc <- ncdf4::nc_open(netcdf.file)

#select the variable
var.name <- names(nc[['var']])[1] #be sure this is the value you want

summary(nc$dim$lon$vals) #if longitude is 0-360 needs to be rotated
tunits <- ncdf4::ncatt_get(nc, "time", "units")
print(tunits)

ncdf4::nc_close(nc)
```

Rotate because the extent is 0-360 not 0-180 (takes a while)
```{r}
dat <- rotate(brick(netcdf.file, varname= var.name))
```

Crop to spatial extent for faster processessing
```{r}
datC <- crop(dat, ext)
```

### Extract Data for x,y point and format using function extract_xy
Cape Grim: 144.689, -40.683 
Mt Read: 145.527, -41.837 
Cradle Mountain: 145.944, -41.637 
```{r}
# Extracts data from raster stack at an x,y location and writes a file for use later named by sitename.
extract_xy <- function(x, y, clim.raster, sitename){
    xy_pt <-	cbind(x, y)
    clim.raster_e <- extract(clim.raster, xy_pt, df = TRUE)
    xy_data <- data.frame(t(clim.raster_e)[-1,])
    colnames(xy_data) <- "O18"
    yr_mo_dy <- substr(row.names(xy_data), 2, 11)
    d <- as.Date(gsub(".", '/', yr_mo_dy, fixed = T))
    xy_data$Date <- d
    xy_data <- xy_data[c("Date", "O18")]
    write.csv(xy_data, paste0("data/", sitename, "_wiso.csv"), row.names=FALSE)
    return(xy_data)
}

wiso_cg <- extract_xy(144.689, -40.683, datC, "CapeGrim")
wiso_mr <- extract_xy(145.527, -41.837, datC, "MtRead")
```

```{r}
plot(wiso_cg$Date, wiso_cg$O18, type="l")

```

### Bring in Cape Grim GNIP Data
```{r read grim data}
cg_dat <- read.csv("data/wiser_gnip-monthly-au-gnipmau01.csv", header=T)[,c("Date", "O18")]
```

```{r}
date_rnge <- range(as.Date(cg_dat$Date))
dates <- as.Date(cg_dat$Date)
wiso_cg_sub <- wiso_cg[as.Date(wiso_cg$Date) %in% dates,]
wiso_mr_sub <- wiso_mr[as.Date(wiso_mr$Date) %in% dates,]
  
sum.tble <- rbind(summary(wiso_cg_sub$O18)[c(1:6)], summary(wiso_mr_sub$O18)[c(1:6)], summary(cg_dat$O18)[c(1:6)])
row.names(sum.tble) <- c("d18O_wiso_cg", "d18O_wiso_mr", "d18O_CapeGrim")
print(sum.tble)
```

Graph wiso and capegrim
```{r}

#gather into single dataframe to allow better plotting
#cg_dat, wiso_cg_sub, wiso_mr_sub, 


ggplot() + 
  geom_line(data = wiso_cg_sub, aes(x = as.Date(Date), y = O18), color = "#00AFBB") + #teal
  geom_line(data = wiso_mr_sub, aes(x = as.Date(Date), y = O18), color = "#FC4E07") + #orange
  geom_line(data = cg_dat, aes(x = as.Date(Date), y = O18), color = "#E7B800") + #yellow
  labs(x='Date',
  y='d18O', 
  color = "Legend") +
    scale_color_manual(values = colors)


```

### Composite Analysis on SAM
Extract Particular Years/Months from datC (climate raster brick)
Create three raster bricks:
datY - a raster brick of all the years and months in the analysis period (set by _years_ variable above).
datSP - a raster brick of the top 10 positive SAM years during the analysis period.
datSN - a raster brick of the top 10 negative SAM years during the analysis period.

datY
```{r}
sel_y <- names(datC)[substr(names(datC), 2, 5) %in% years] 
#use that index to subset the raster brick
datY <- subset(datC, sel_y, value=T)
tail(names(datY), 30) 
```


Read in Marshall SAM data and subset to period, extract 90th percentile highest/lowest
```{r}
sam <- read.table("data/marshallSamAnn.txt")
samY <- sam[row.names(sam) %in% years, ]
spN <- row.names(samY[samY$ANN < (quantile(samY$ANN, 0.1)),]) #bottom 10% annual SAM
spP <- row.names(samY[samY$ANN > (quantile(samY$ANN, 0.9)),]) #top 10% annual SAM
```

Read in Monthly Marshall SAM data 
```{r}
sam_M <- read.table("data/marshallSamMon.txt")
```

Make into growing season years
```{r}

m <- lapply(sam_M, fun=mean, na.rm = TRUE)

```

```{r}

```


Although - SAM seasons - which year?
### Better to look at the growing season, rather than calendar year
First create a factor for growth.years where growth.year begins in Oct previous year and runs through Sept of same year.
Adjust if you just want Oct thru Feb
mkann (in the seas package) not necessarily needed
Need an index/factor that takes all layers (Z) and applies a conversion to say what year the month needs to be applied to (Oct 2015 -> 2016 or other way around). Can probably create this sequence on my own
```{r}
st.mon <- "10"  
yr_mo_dy <- substr(names(datY), 2, 11)
d <- as.Date(gsub(".", '/', yr_mo_dy, fixed = T)) 
start <- paste0(min(substr(names(datY), 2, 5)),"-", st.mon, "-15")
grow.years <- mkann(d, as.Date(start)) #an index of the growing season years.  note first and last are incomplete as is last year.
```

use this factor to create means by growth.year
```{r}
 s <- stackApply(datY, indices=grow.years, fun=mean, na.rm=TRUE) #adjust names of layers to match spN
```


#####NEEDS UPDATE FROM HERE TO MATCH s 
datSN - negative SAM years only
```{r}
sel_spN <- names(datY)[substr(names(datY), 2, 5) %in% spN] 
#use that index to subset the raster brick
datSN <- subset(datY, sel_spN, value=T)
tail(names(datSN), 30) 
dim(datSN) #should have 120 layers
```

datSP - positive SAM years only
```{r}
sel_spP <- names(datY)[substr(names(datY), 2, 5) %in% spP] 
#use that index to subset the raster brick
datSP <- subset(datC, sel_spP, value=T)
tail(names(datSP), 30) 
dim(datSP) #should have 120 layers
```

### Calculate a composite (subtract the datSP from the mean conditions datY)
Climatological mean:
```{r}
datM <- calc(datY, mean)
```
Mean of positive SAM years
```{r}
datSP.mean <- calc(datSP, mean)
```
Mean of negative SAM years
```{r}
datSN.mean <- calc(datSN, mean)
```

Calculate the anomaly = Observed Value - Mean Value
Alternates?: raster, calc, stackapply
Calc more efficient over whole data
stackapply if you only use some of the layers, pass a factor through the calculation
```{r}
P.anom <- datSP.mean - datM
N.anom <- datSN.mean - datM
```
Next steps: bootstrapping

Combine the two layers into a brick for easier plotting using levelplot
```{r}
anom.list <- list(P.anom, N.anom) 
anom.brick <- raster::brick(anom.list)
names(anom.brick) <- c("P.anom", "N.anom")
```

Plot
```{r}
coasts <- ne_coastline(scale=50, "sp")
coasts <- crop(coasts, ext)
levelplot(anom.brick, colorkey=list(title="Pos, Neg Anom", space="bottom"), margin=TRUE, 
          names.attr=names(anom.brick),     par.settings=rasterTheme(viridis_pal(option = "D")(255))) + 
   layer(sp.polygons(coasts, lwd=1)) 
```

